---
title: "Getting Started with jinx"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with jinx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

jinx is the R-Ladies GitHub organization management bot.
It is an R package deployed via GitHub Actions with a registered
GitHub App identity (`jinx[bot]`).

## Architecture

```
Issue comment (/jinx invite ...)
  -> issue_comment workflow triggers
    -> installs jinx R package
      -> calls jinx::parse_command() + jinx::execute_command()
        -> gh::gh() API calls as jinx[bot]
          -> reply comment posted
```

jinx is a standard R package. GitHub Actions workflows call exported R
functions. The `gh` package handles GitHub API authentication via the
`GITHUB_TOKEN` environment variable. When using a GitHub App token (via
`actions/create-github-app-token`), API calls and comments appear as
`jinx[bot]`.

## Prerequisites

### 1. Register the GitHub App

1. Go to **Settings > Developer settings > GitHub Apps > New** for your org
2. Name: `jinx`
3. Permissions:
   - Repository: Issues (R/W), Pull Requests (R/W), Contents (Read)
   - Organization: Members (R/W), Administration (Read)
4. Events: `issue_comment`, `issues`, `pull_request`
5. Install on the R-Ladies org

### 2. Configure secrets

In the jinx repo (or org-wide):

- **Variable** `JINX_APP_ID` -- the GitHub App's ID
- **Secret** `JINX_PRIVATE_KEY` -- the GitHub App's private key

Additional secrets for specific modules:

| Secret | Module |
|--------|--------|
| `AIRTABLE_API_KEY` | Airtable sync |
| `BSKY_USER`, `BSKY_PASS` | Bluesky announcements |
| `MASTODON_TOKEN` | Mastodon announcements |
| `LINKEDIN_ACCESS_TOKEN` | LinkedIn announcements |
| `SLACK_TOKEN` | Slack management |
| `MAILCHIMP_API_KEY` | Newsletter |

### 3. Installation

```{r}
#| eval: false
remotes::install_github("rladies/jinx")
```

## The `/jinx` command system

jinx responds to commands posted as issue comments in any repo where it
is installed. Commands start with `/jinx` followed by an action:
```
/jinx invite @username to website
/jinx offboard @username from blog
/jinx report weekly
/jinx chapter-setup Berlin Germany
/jinx help
```

Commands are parsed by `parse_command()` and dispatched by
`execute_command()`:

```{r}
#| eval: false
library(jinx)

cmd <- parse_command("/jinx invite @octocat to website")
cmd
#> $action
#> [1] "invite"
#> $username
#> [1] "octocat"
#> $team
#> [1] "website"
```

## Modules

jinx is organized into modules, each handling a specific domain:

| Module | Functions |
|--------|-----------|
| Team management | `gt_invite()`, `gt_create_offboarding()`, `gt_finalize_onboarding()` |
| Announcements | `announce_post()`, `post_bluesky()`, `post_mastodon()` |
| Directory | `validate_directory_pr()`, `crop_directory_image()` |
| Blogs | `create_blog_entry()`, `check_blog_links()` |
| Chapters | `create_chapter()`, `check_chapter_health()` |
| Reports | `generate_report()`, `publish_report()` |
| PR review | `review_pr()`, `check_pr_naming()` |
| Airtable sync | `sync_directory_airtable()`, `sync_global_team_airtable()` |
| Website | `auto_merge_pending()`, `greet_contributor()` |
| Chapter monitoring | `monitor_chapter_status()`, `prepare_inactivity_emails()` |
| Slack | `send_slack_invites()`, `subscribe_slack_rss()` |
| GHA dashboard | `generate_gha_dashboard()`, `publish_gha_dashboard()` |
| Contributors | `list_contributors()`, `update_contributors_list()` |

## Configuration

jinx loads configuration from YAML files in `inst/config/`:

```{r}
#| eval: false
config <- load_teams_config()
names(config$teams)
#>  [1] "abstract-review"    "blog"               "campaigns"
#>  [4] "chapter-activity"   "chapter-onboarding" "coc"
#>  [7] "communications"     "community-slack"    "conference-liaison"
#> [10] "directory"          "meetup-pro"         "mentoring"
#> [13] "rocur"              "translation"        "website"
```

- `teams.yml` -- team definitions with roles, repos, notification channels
- `review-rules.yml` -- PR review assignment and labeling rules
- `labels.yml` -- file path to label mappings

## Template system

Templates live in `inst/templates/` as markdown files with `<KEY>`
placeholders. Team-specific extras live in `inst/templates/teams/`:

```
inst/templates/
  global-team-onboarding.md      # Base onboarding template
  global-team-offboarding.md     # Base offboarding template
  chapter-setup.md               # Chapter setup checklist
  teams/
    website.md                   # Website team extras
    blog.md                      # Blog team extras
    ...
```

Templates are rendered via `render_template()` which replaces `<KEY>`
placeholders with values from a named list.

## Cross-repo PR review

Add this workflow to any R-Ladies repo for automated PR review:

```yaml
# .github/workflows/jinx-review.yml
name: PR Review
on:
  pull_request:
    types: [opened, reopened, synchronize]
jobs:
  review:
    uses: rladies/jinx/.github/workflows/pr-review.yml@main
    secrets: inherit
```

## Next steps

See `vignette("workflows")` for step-by-step guides for each admin
workflow.
