---
title: "Workflow Reference for R-Ladies Org Admins"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow Reference for R-Ladies Org Admins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette walks through each admin workflow jinx supports.
For setup and architecture details, see `vignette("getting-started")`.

## Onboarding a new team member

**Via issue comment:**
```
/jinx invite @username to website
```

**What happens:**

1. `gt_invite()` sends an org invitation to the user
2. A tracking issue is created with the team-specific onboarding checklist
3. Once the user accepts, `gt_finalize_onboarding()` adds them to the
   appropriate team(s) and repos
4. The tracking issue is closed

**Programmatically:**
```{r}
#| eval: false
gt_invite("octocat", "website")
```

**Checking pending invitations:**
```{r}
#| eval: false
gt_check_invitations()
```

## Offboarding a team member

**Via issue comment:**
```
/jinx offboard @username from blog
```

**What happens:**

1. `gt_create_offboarding()` opens an offboarding tracking issue
2. The issue lists cleanup steps (remove from team, repos, external services)
3. Once complete, `gt_finalize_offboarding()` removes org access and closes
   the issue

**Programmatically:**
```{r}
#| eval: false
gt_create_offboarding("octocat", "blog")
```

## Setting up a new chapter

**Via issue comment:**
```
/jinx chapter-setup Berlin Germany
```

**What happens:**

1. `chapter_create_setup()` opens a setup tracking issue
2. The issue contains a checklist: create repos, add organizers, configure
   Meetup, set up social media
3. `create_chapter()` generates the chapter JSON for the directory
4. `create_chapter_json_pr()` opens a PR to add the chapter to the directory

**Updating an existing chapter:**
```
/jinx chapter-update Berlin Germany
```

## Monitoring chapter health

**Scheduled workflow** runs periodically via `chapter_check_health()`.

**Via issue comment:**
```
/jinx chapter-health
```

**What happens:**

1. Reads chapter activity data (commits, issues, events)
2. Classifies chapters as active, quiet, or inactive
3. Posts a summary with recommendations

**Outreach to inactive chapters:**
```{r}
#| eval: false
statuses <- monitor_chapter_status()
prepare_inactivity_emails(statuses)
```

## Blog post workflow

### Adding a new blog entry

**Via issue comment:**
```
/jinx blog-add https://example.com/my-rladies-post
```

**What happens:**

1. `blog_create_entry()` fetches metadata from the URL
2. Creates a YAML entry and opens a PR to the blog repo
3. `blog_post_checklist()` adds a review checklist to the PR

### Checking blog links

```
/jinx blog-check-links
```

Runs `blog_check_links()` to find broken URLs across all blog entries.

### Auto-merge on publish date

`website_merge_pending()` runs on a schedule and merges blog PRs whose
publish date has arrived. This is triggered by the
`website-merge-pending.yml` workflow.

## Announcing blog posts

After a blog post merges, announce it across platforms:

**Via issue comment:**
```
/jinx announce https://rladies.org/blog/my-post/
```

**What happens:**

1. `announce_post()` reads the post's YAML front matter
2. `create_announcement_message()` builds the message with hashtags
3. Posts to Bluesky (`post_bluesky()`), Mastodon (`post_mastodon()`),
   LinkedIn (`li_post_write()`), and newsletter (`send_newsletter()`)

Each platform is posted independently -- failures on one do not block others.

## Running reports

### Activity reports

**Via issue comment:**
```
/jinx report weekly
/jinx report monthly
```

**Programmatically:**
```{r}
#| eval: false
report <- report_generate(type = "weekly")
report_publish(report)
```

### Chapter health reports

```
/jinx report chapters
```

```{r}
#| eval: false
url <- chapter_report_health()
```

## Managing the GHA dashboard

```
/jinx gha-dashboard
```

**What happens:**

1. `gha_generate_dashboard()` queries workflow run status across all org repos
2. `gha_publish_dashboard()` posts a formatted status table as a GitHub issue

**Scheduled workflow** runs weekly via `gha-dashboard.yml`.

## Airtable sync

**Scheduled workflow** runs weekly via `airtable-sync.yml`.

**Programmatically:**
```{r}
#| eval: false
sync_directory_airtable()
sync_gt_airtable()
```

Syncs directory entries and global team data between Airtable and GitHub.
Creates PRs for any changes detected.

## Slack onboarding

```{r}
#| eval: false
slack_invites_send(dry_run = TRUE)
```

Fetches pending invitees from Airtable and sends Slack workspace
invitations. Use `dry_run = TRUE` (the default) to preview before sending.

**RSS feed integration:**
```{r}
#| eval: false
subscribe_slack_rss()
```

Subscribes Slack channels to the R-Ladies blog RSS feed.

## Directory maintenance

**Validating directory PRs:**

The `validate_directory_pr()` function runs automatically on PRs that
modify directory entries. It checks:

- YAML schema compliance
- Filename conventions (`validate_entry_filename()`)
- Image optimization (`crop_directory_image()`, `optimize_image()`)
- Social handle verification (`verify_social_handles()`)

## Contributor recognition

**List contributors for a repo:**
```
/jinx contributors jinx
```

**Update the contributors list:**
```
/jinx contributors update jinx
```

**Org-wide contributor stats:**
```
/jinx contributors org
```

**Automated greetings:**

- `contributor_welcome()` posts a welcome comment on first-time PRs
- `contributor_thank()` posts a thank-you when PRs are merged

Both are triggered by the `welcome-contributor.yml` and
`thank-contributor.yml` workflows.

## PR review automation

PR review runs automatically on pull requests via reusable workflows.

**What happens:**

1. `review_pr()` applies the rules from `review-rules.yml`
2. Reviewers are assigned based on file paths and team ownership
3. Labels are applied based on `labels.yml` mappings
4. `check_pr_naming()` enforces branch and title conventions

**See** `vignette("getting-started")` for how to add PR review to other
repos.

## Stale issue reminders

```
/jinx remind stale
```

`gt_remind_stale()` scans for onboarding/offboarding issues that have
gone stale and posts reminder comments.
